---
title: "Sales Trends"
author: "Cassandra Bayer"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Controller.R")
```

# Initial Processing
## Load data, clean it up, and do some brief analytics
As a first pass, I load the data, drop the null columns, extra the month and the month name, and do a quick glance of the data. I see that the classes of the data were passed into R as intended, and the distribution of the data seems fairly normal.
```{r}
orders <- data.table(read.xlsx2(file = "XLS_takehome_NA.xlsx", 
                                sheetName = "Orders",
                                stringsAsFactors = F,
                                colClasses = c("Date", "Date", "character", "numeric")))

## clean up data
orders <- orders[, 1:4][order(dateordered)]
orders[, month := month(dateordered)]
orders$monthName <- month.name[ orders$month]

## descriptives
sapply(orders, class)
summary(orders)
```

# Aggregation I
At first glance, one may think that the rate of returns is rising-- the sheer number of returns has increased over time. However in order to answer the first question, I did an aggregation to get the return rate by month. I did so by calculating the number of orders returned per the number of orders placed by the month in which that order was placed. According to these trends, **the return rate has been steadily falling since August.**
<br><br>
I believe that the return rate calculated here is indicative of how the company was performing or the quality of the mattress at the the time of sale.
The plot, despite being out of order, shows the return rates over time. 
```{r warning=FALSE}
returnRate <- orders[, .( orders = sum(orders[orderstatus == "complete"], na.rm = T),
                          returns = sum(orders[orderstatus == "returned"], na.rm = T),
                          returnRate = sum(orders[orderstatus == "returned"], na.rm = T)/
                           sum(orders[orderstatus == "complete"], na.rm = T)),
                     by = .(monthName, month)]
## clean up
returnRate[, returnRate := round(returnRate,2)]
print(returnRate)

## plot
plot_ly(data = returnRate, 
        x = ~monthName, 
        y =  ~returnRate, 
        mode = 'lines', 
        colors = "Set1")

```

# Aggregation II

As a gut check, I wanted to aggregate the data in a different way. I aggregated the data by the date of return, which shows the sheer volume of mattresses being sold and returned by month, regardless of their sale date. While I believed this calculation to be less valuable, I created it to rest assured. Again, we can see that, although the number of returns is rising, the rate is not as telling when ignoring the month in which the order was placed. 
```{r warning=FALSE}
returnsByMonth <- orders[, .(returns = sum(orders[orderstatus == "returned"], na.rm = T)),
                         by = monthName]

ordersByMonth <- orders[, .(orders = sum(orders[orderstatus == "complete"], na.rm = T)),
                        by = monthName]
## join
setkey(returnsByMonth, monthName)
setkey(ordersByMonth, monthName)
returnsByMonth <- ordersByMonth[returnsByMonth]

## clean up
returnsByMonth[!is.na(monthName), returnRate := returns/orders, by = monthName]
returnsByMonth <- returnsByMonth[c(1, 5, 4, 3, 2)]
print(returnsByMonth)

## plot
plot_ly(returnsByMonth, x = ~monthName) %>%
  add_trace(y = ~orders, name = 'Orders',mode = 'lines') %>%
  add_trace(y = ~returns, name = 'Returns', mode = 'lines') 
```

# Analysis
**Some examples of factors highlighted by the first aggregation include:** <br>
- Changing (bettering) quality of product over time <br>
- Changing (bettering) customer experience at the point of sale (whether customers are being matched to the right product (mattress) at the point of sale) 
<br><br>
**Some examples of factors highlighted by the second aggregation include:**<br>
- Changes to consumer behavior in reaction to new competitors<br>
- Changes to post-sale responsiveness of customer support<br>
- Other factors not associated with date of point-of-sale

# SQL Query
Using a package in R, I created and ran a SQL query to get the number of completed orders by date. If the audience also wanted to see an aggregation by month, I would add either a `MONTH` or a `DATEPART` to the query.
```{r}
casper_orders <- orders

sqldf("SELECT SUM(orders) as totalOrders, dateordered 
      FROM casper_orders 
      WHERE orderstatus = 'complete' 
      GROUP BY dateordered;")
```

